Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Please @JJ explain better the objective and the measures we are going to use in the abstract
Minor English corrections in the intro

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to fully harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
evolutionary computation experiments with different implementations of
a volunteer computing framework called {\sf NodIO}, designed to be
easily deployable in using free cloud resources. We use  
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything in their machines, they just
need to open a web page  
in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX are. Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system
independence. This system, however, is not intended to substitute
frameworks running in dedicated hardware or the cloud, but to
complement it. In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed.

@@

In this paper we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments using a
Platform as a Service, and a simple Ajax client that does not
restart. Second, we change a bit the client so that it uses web
workers, that is, it runs in the background and it restarts. Third, we use the same
client but with a harder problem that cannot be solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. With these three experiments we want mainly to try and model
the behavior of users and how different problems and implementations
have an influence on it.

>>>>

Using the browser as a computing resource in volunteer computing presents several advantages, 
but it remains a challenge to fully harness the browser's capabilities and
to model the user's behavior so that those capabilities can be
leveraged optimally. In this paper, we present the results of several
evolutionary computation experiments with different implementations of
a volunteer computing framework called {\sf NodIO}, designed to be
easily deployable in using free cloud resources. We use  
%Pablo: "deployable in using free cloud resources" is correct?
% why not? - JJ
different implementations to find out which one is 
able to get the user to lend more computing cycles and test different
problems to check the influence it has on said performance.
% Pablo: so the objective is "to determine which browser conf. is able
% to get more computing power from the users", it isn't? 
% (to keep this in mind the rest of the paper). What is a "browser configuration"?
% (explain it with a phrase in the abstract)
% No, we don't. Please read the paper!!!! - JJ
% FERGU: Sorry, JJ, but the objective is not clear to me yet here, 
% even if the implementation comparison I wrote is in the conclusions as the first 
% objective of the paper. Also, the metrics used should be mentioned in the abstract.
% For example: "The first objective of this paper is to compare different implementations 
% of NodIO (with and without Web Workers), measuring the time spent by the users. 
% The second objective is to model the user behavior by comparing whatever".
% Also, the two compared implementations (w and w/ Web Worker) should be mentioned.
From these experiments we can already draw some conclusions, besides
the fact that volunteer computing can be a valuable computing
resource: the user has to be kept engaged in some way to be able to
use as many computing cycles as possible, the client has to be built
to use the computer capabilities fully, and, finally, the user
contributions follow a common statistical distribution. It is also
essential to be as open as possible with software and data.

@@

%PABLO: He añadido el siguiente párrafo para el issue #31
There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript. However, one of the advantages of our architecture is that
clients do not need to install anything on their machines, they just
%Pablo: According English Stack Exchange, you install "ON" a machine (but "IN" a directory)
need to open a web page  
in a compatible, that is, almost any, browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Another advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX are. Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system
independence. This system, however, is not intended to substitute
frameworks running on dedicated hardware or the cloud, but to
%Pablo: same here, I learnt this with my paper of EAs "on clusters"
complement it. In general, several volunteers will be needed to
overcome the overhead involved in using a client-server architecture
and this number of volunteers is obviously not guaranteed.

@@

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments using a
Platform as a Service, and a simple Ajax client that does not
restart. Second, we change a bit the client so that it uses web
workers, that is, it runs in the background and it restarts. Third, we use the same
client, but with a harder problem that cannot be solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. With these three experiments we want mainly to try and model
the behavior of users and how different problems and implementations
have an influence on it.
